SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-09-09 maintex.tex
updated the radial profile section

\subsection{Emission Line Extraction}

%%% SPFIT
We use {\sc SPFIT} to get measurements of emission line fluxes, equivalent widths, velocities, and velocity dispersions of 19 emission lines and measurements of stellar masses, ages, [Fe/H], and the kinematics of the stellar populations.

@@

The V-band extinction is then used to calculate the wavelength dependent extinciton A$_{\lambda}$;

\begin{equation}
A_{\lambda}(x) = A_V \left(a(x) + b(x) / R_v \right)
\end{equation}

where $x = 1/\lambda$ and $a(x)$ and $b(x)$ are the coefficients for the extinction curve from \citet{ODonnell:1994}. The extinction corrected flux is then;

\begin{equation}
F'(\lambda) = F(\lambda) * 10^{0.4 A_{\lambda}(x)}
\end{equation}

This extinction correction is applied to all of the extracted emission lines.

\subsection{Star Formation Rate}

We measure the SFR from the H$\alpha$ luminosity, L$_{{\rm H}\alpha}$, calculated from the H$\alpha$ flux.  We use the SFR formula, Equation \ref{eq:sfr}, from \citet{Murphy:2011} which uses a Kroupa IMF, Solar metallicity, a constant SFR at an age of 100 Myr, and Case-B recombination.

@@

Since the stellar mass of a galaxy is not uniformly distributed within the galaxy, we normalize the SFR by the local stellar mass, giving us the specific star formation rate (sSFR). The local stellar masses used here is the stellar mass calculated by {\sc SPFIT}'s fit to the stellar continuum within each spaxel (spatial pixel).

@@

This will allow us to compare the SFR in high mass regions, like the centers of the galaxies, to the SFR lower mass regions, like the disk of the galaxies. We have also found a strong correlation between sSFR and \ewha\ and have found the same results using \ewha. This is useful to check as \ewha\ is an observable and independent of {\sc SPFIT}'s measurement of the stellar mass.

%We use the BPT diagnostic \citep{Baldwin:1981} to remove spaxels which were not ionized by star formation. To separate ionization sources, we use the theoretical maximum starburst line from \citet{Kewley:2001} (Equation \ref{eq:kewley}).

%\begin{equation}\label{eq:kewley}
%\rm{log\Big(\frac{[O\,\textsc{iii}]\lambda 5007}{H\beta}\Big) = \frac{0.61}{log([N\,\textsc{ii}]\lambda 6584 / H\alpha) - 0.47} + 1.19}
%\end{equation}

%Where \OIII$\lambda$5007, H$\beta$, \NII$\lambda$6584, and H$\alpha$ are emission line fluxes within each MaNGA spaxel. We use the \citet{Kewley:2001} demarcation line as the limit for classifying star formation in the MaNGA spaxels.

It has been shown that hot low mass evolved stars (HOLMES) in retired galaxies can mimic H\,{\sc ii} regions and AGN on the BPT diagram \citep{Stasinska:2008}. To mask retired spaxels, we implement a H$\alpha$ equivalent width (EW) cut of \ewha\ $>$ 3\AA\ \citep{Cid-Fernandes:2011}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Geometry of Galaxies}\label{sec:radial}

@@

In order to spatially characterize the star formation in the paired galaxies we build radial profiles of sSFR. First, the geometry of the galaxies need to be defined. Specifically we will need position angle, inclination angle, and the effective radius of each of the MaNGA targets. We use the $r$-band elliptical Petrosian apertures from the NSA catalog and the $r$-band S\'ersic apertures from \citet{Simard:2011} to define the geometries of the galaxies.

The NSA catalog has complete coverage over the MaNGA sample (since MaNGA selects its targets from this catalog); however, it tends to fail to properly fit paired galaxies with close on-sky separations. We found that the apertures from \citet{Simard:2011} work better for these close galaxies, but the catalog does not completely cover the MaNGA sample. We use NSA catalog for the outside-IFU pair sample because they are well separated on the sky and the \citet{Simard:2011} catalog for the inside-IFU sample. In the cases where an Inside-IFU paired galaxy is not covered by \citet{Simard:2011}, we use the ellipse from the NSA catalog.

We fit the geometry with aperture from both catalog when available. When using the first method of pair-control comparison, where pairs and controls are grouped into stellar mass bins, we fit the control galaxies with the NSA apertures. When making the comparison with the second method, paired galaxies fitted with the NSA aperture are compared against controls fitted with the NSA aperture and paired galaxies fitted with the \citet{Simard:2011} apertures are compared against controls fitted with the \citet{Simard:2011} apertures.

@@

Where $q$ is the intrinsic oblateness. We use the empirically determined oblateness of $q = 0.13$, from \citet{Giovanelli:1994}.

The inclination angle, along with the galaxy's position angle, is used to deproject the geometries of the galaxies. We use the 50\% half light radius, the effective radius (\reff) to scale the sizes of the galaxies. Doing this will allow use the compare galaxies of different sizes against each other.

\subsection{Radial Profiles}

@@

We use our catalog of photometric objects and the \citet{Simard:2011} catalog to mask foreground/background objects within the MaNGA IFUs which may contaminate the data. Objects in our photometric catalog are masked out with a 2\arcsec\ circular aperture and the galaxies from \citet{Simard:2011} are masked out with a 2.0 R/\reff\ elliptical aperture, using the r-band S\'ersic apertures from the catalog.

>>>>

\subsection{Specific Star Formation Rate}

%%% SPFIT
We use our spectral fitting code, {\sc spfit}, to get measurements of emission line fluxes, equivalent widths, velocities, and velocity dispersions of 19 emission lines and measurements of stellar masses, ages, [Fe/H], and the kinematics of the stellar populations.

@@

We measure the SFR from the extinction corrected H$\alpha$ luminosity, L$_{{\rm H}\alpha}$, calculated from the H$\alpha$ flux.  We use the SFR formula, Equation \ref{eq:sfr}, from \citet{Murphy:2011} which uses a Kroupa IMF, Solar metallicity, a constant SFR at an age of 100 Myr, and Case-B recombination.

@@

Since the stellar mass of a galaxy is not uniformly distributed within the galaxy, we normalize the SFR by the local stellar mass, $M_*$, giving us the specific star formation rate (sSFR). The local stellar masses used here is the stellar mass calculated by {\sc SPFIT}'s fit to the stellar continuum within each spaxel (spatial pixel).

@@

This will allow us to compare regions to high mass surface density to regions of low mass surface density.

We check the results of the radial profiles of this sSFR against the radial profiles of \ewha\ and find that they are in good agreement. \ewha\ is an observable while our measurement of the sSFR is dependent on both the H$\alpha$ luminosity and {\sc spfit}'s calculation of the stellar mass, so this check helps confirm the quality of our sSFR measurement.

%It has been shown that hot low mass evolved stars (HOLMES) in retired galaxies can mimic H\,{\sc ii} regions and AGN on the BPT diagram \citep{Stasinska:2008}. To mask retired spaxels, we implement a H$\alpha$ equivalent width (EW) cut of \ewha\ $>$ 3\AA\ \citep{Cid-Fernandes:2011}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Radial Profiles}\label{sec:radial}

@@

In order to spatially characterize the star formation in the paired galaxies we build radial profiles of sSFR. First, the geometry of the galaxies needs to be defined. Specifically we will need position angle, inclination angle, and the effective radius of each of the MaNGA targets. We use the $r$-band elliptical Petrosian apertures from the NSA catalog and the $r$-band S\'ersic apertures from \citet{Simard:2011} to define the geometries of the galaxies.

The NSA catalog has complete coverage over the MaNGA sample (since MaNGA selects its targets from this catalog); however, it tends to fail to properly fit paired galaxies with close on-sky separations. We found that the apertures from \citet{Simard:2011} work better for these close paired galaxies; however, the catalog does not completely cover the MaNGA sample. We use the NSA catalog for the outside-IFU pair sample because they are well separated on the sky. The \citet{Simard:2011} catalog is used for the inside-IFU sample. If the paired galaxy is not covered by \citet{Simard:2011}, we use the ellipse from the NSA catalog.

We fit the geometry with apertures from both catalog when available. When using the first method of pair-control comparison, where pairs and controls are grouped into stellar mass bins, we fit the control galaxies with the NSA apertures. When making the comparison with the second method, paired galaxies fitted with the NSA aperture are compared against controls fitted with the NSA aperture and paired galaxies fitted with the \citet{Simard:2011} apertures are compared against controls fitted with the \citet{Simard:2011} apertures.

@@

Where $b/a$ is the major-to-minor axis ratio and $q$ is the intrinsic oblateness. We use the empirically determined oblateness of $q = 0.13$, from \citet{Giovanelli:1994}.

The inclination angle, along with the galaxy's position angle, is used to deproject the geometries of the galaxies. We use the 50\% half light radius (i.e. the effective radius, \reff) to scale the sizes of the galaxies. Doing this will allow us to compare galaxies of different sizes against each other.

@@

We use our catalog of photometric objects and the \citet{Simard:2011} catalog to mask foreground/background objects within the MaNGA IFUs which may contaminate the data. Objects in our photometric catalog are masked out with a 2\arcsec\ circular aperture and the galaxies from \citet{Simard:2011} are masked out with a 2.0 R/\reff\ elliptical aperture, using the r-band S\'ersic apertures from the catalog. This gives us an individual radial profile of the sSFR for all paired and control galaxies.
